---
title: "glmmExample"
author: "Sara Stoudt"
date: "April 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generalized Linear Mixed Model (GLMM) example

Now we will explore a real-world example and show how to change some of the more nitty-gritty details. These are  

```{r}
library("MCMCmaxlik")
```

Here we set up the salamander data.

```{r}
require(glmm)
data(salamander)
names(salamander)
nrow(salamander) 

isRR=ifelse(salamander$Cross=="R/R",1,0)
#head(isRR)

isRW=ifelse(salamander$Cross=="R/W",1,0)
#head(isRW)

isWR=ifelse(salamander$Cross=="W/R",1,0)
#head(isWR,20)

isWW=ifelse(salamander$Cross=="W/W",1,0)
#head(isWW,20)


```

Here we set up the model.

```{r model}
glmmCode <- nimbleCode({
  for(i in 1:N){
  logit(theta[i])<-beta1*isRR[i]+beta2*isRW[i]+beta3*isWR[i]+beta4*isWW[i]+REF[i]+REM[i]
  y[i]~dbin(theta[i],1)
  REF[i]~dnorm(0,varF)
  REM[i]~dnorm(0,varM)
  }
  
  
  beta1 ~ dnorm(0, sd = 10000)
  beta2 ~ dnorm(0, sd = 10000)
  beta3 ~ dnorm(0, sd = 10000)
  beta4 ~ dnorm(0, sd = 10000)
  varF ~ dunif(0, 1000)
  varM ~ dunif(0, 1000)
  
})


glmmConstants <- list(N = nrow(salamander))

glmmData <- list(
  isRR=isRR,isWR=isWR,isRW=isRW,isWW=isWW,y=salamander$Mate
)

glmmInits <- list(beta1 = 1, beta2 = 1, beta3=1,beta4=1,varF=5,varM=5)

glmmMod <- nimbleModel(code=glmmCode, constants=glmmConstants, data=glmmData, inits=glmmInits, check = FALSE)

glmmParamNodes <- glmmMod$getNodeNames(topOnly = T,stochOnly=T)

Cglmm <- compileNimble(glmmMod)

```


## Running the algorithms

First, we compile the algorithms in NIMBLE using `buildMCMCmaxlik()`.

```{r compile}
compiledFunsGlmm <- buildMCMCmaxlik(glmmMod, glmmParamNodes)
```

Then we can run the algorithms using `computeMLE()`. For Adadelta and Adam we use the default settings used by the creators of these step size schedules, but we do allow the user to change these as desired. We want to give flexibility for those who have experience with these algorithms in other contexts and may want to customize our impelmentations to fit their needs. 

Default Parameters for Adagrad: delta = 1e-04, eps = 1e-2

Default Paramters for Adam: beta1=0.9, beta2=0.999

```{r algorithm1D, results="hide"}
resultsGLMMadadeltaDefault <- computeMLE(model=logreg, 
                                    paramNodes = paramNodesLogreg, 
                                    paramInit=c(10, 10),
                                    compiledFuns=compiledFunsLogReg,
                                    method="adadelta")

# Warning: in nimArr_2_ManyModelAccess, accessor larger than NimArr!
# |-------------|-------------|-------------|-------------|
# |-------------------------------------------------------|
# longer object length is not a multiple of shorter object lengthlonger object length is not a multiple of shorter object lengthlonger object length is not a multiple of shorter object lengthlonger object length is not a multiple of shorter object length[1] -0.3131166 -0.3095830  0.1766880
# [1]  9.686883  9.690417 10.176688
# longer object length is not a multiple of shorter object lengthError in paramMatrix[iter + 1, ] <- thetaNew : 
#   number of items to replace is not a multiple of replacement length
```


```{r algorithm2D, results="hide"}
resultsGLMMadamDefault<- computeMLE(model=logreg, 
                                    paramNodes = paramNodesLogreg, 
                                    paramInit=c(10, 10),
                                    compiledFuns=compiledFunsLogReg,
                                    method="adam")

# Warning: in nimArr_2_ManyModelAccess, accessor larger than NimArr!
# longer object length is not a multiple of shorter object lengthlonger object length is not a multiple of shorter object lengthlonger object length is not a multiple of shorter object lengthError in paramMatrix[iter + 1, ] <- thetaNew : 
#   number of items to replace is not a multiple of replacement length
```

Now let's change some parameters.

```{r algorithm1, results="hide"}
resultsGLMMadadelta <- computeMLE(model=logreg, 
                                    paramNodes = paramNodesLogreg, 
                                    paramInit=c(10, 10),
                                    compiledFuns=compiledFunsLogReg,
                                    method="adadelta",delta=1e-03,eps=.1)

# Warning: in nimArr_2_ManyModelAccess, accessor larger than NimArr!
# |-------------|-------------|-------------|-------------|
# |-------------------------------------------------------|
# longer object length is not a multiple of shorter object lengthlonger object length is not a multiple of shorter object lengthlonger object length is not a multiple of shorter object lengthlonger object length is not a multiple of shorter object length[1] -0.9392501 -0.8957062  0.2105871
# [1]  9.060750  9.104294 10.210587
# longer object length is not a multiple of shorter object lengthError in paramMatrix[iter + 1, ] <- thetaNew : 
#   number of items to replace is not a multiple of replacement length
```


```{r algorithm2, results="hide"}
resultsGLMMadam<- computeMLE(model=logreg, 
                                    paramNodes = paramNodesLogreg, 
                                    paramInit=c(10, 10),
                                    compiledFuns=compiledFunsLogReg,
                                    method="adam",beta1=.8,beta2=0.99)

# Warning: in nimArr_2_ManyModelAccess, accessor larger than NimArr!
# longer object length is not a multiple of shorter object lengthlonger object length is not a multiple of shorter object lengthlonger object length is not a multiple of shorter object lengthError in paramMatrix[iter + 1, ] <- thetaNew : 
#   number of items to replace is not a multiple of replacement length
```

How do we do?

```{r 1DResult}
resultsGLMMadadeltaDefault$MLE
resultsGLMMadamDefault$MLE
resultsGLMMadadelta$MLE
resultsGLMMadam$MLE
```

