---
title: "Example: pump"
author: "Johnny Hong"
date: "March 29, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pump example

As a simple example, we use the Gamma-Poisson hierachical model for nuclear pumps.

```{r}
library("MCMCmaxlik")
```

NIMBLE uses the BUGS language to specify a hierarchical model.
```{r model}
pumpCode <- nimbleCode({
  for (i in 1:N) {
    theta[i] ~ dgamma(alpha, beta)
    lambda[i] <- theta[i] * t[i]
    x[i] ~ dpois(lambda[i])
  }
  alpha ~ dexp(1.0)
  beta ~ dgamma(0.1, 1.0)
})
pumpConsts <- list(N=10,
                   t=c(94.3, 15.7, 62.9, 126, 5.24,
                         31.4, 1.05, 1.05, 2.1, 10.5))
pumpData <- list(x=c(5, 1, 5, 14, 3, 19, 1, 1, 4, 22))
pumpInits <- list(alpha=1, beta=1,
                  theta=rep(0.1, pumpConsts$N))

pump <- nimbleModel(code=pumpCode, name='pump', constants=pumpConsts,
                    data=pumpData, inits=pumpInits)
paramNodesPump <- pump$getNodeNames(topOnly=T)
Cpump <- compileNimble(pump)
```


## Running the algorithms
First, we compile the algorithms in NIMBLE using `buildMCMCmaxlik()`.
```{r compile}
compiledFunsPump <- buildMCMCmaxlik(pump, paramNodesPump)
```

Then we can run the algorithms using `computeMLE()`.
```{r adam, results="hide"}
resultsPumpAdam <- computeMLE(model=pump, 
                              paramNodes = paramNodesPump, 
                              paramInit=c(10, 10),
                              compiledFuns=compiledFunsPump)
```

The output contains the history of iterates, the number of iterations and the 20 percent
trimmed average of the last 20 iterates.
```{r adamResult}
resultsPumpAdam$MLE
```

The history of iterates allows the users to study the convergence of the algorithms:
```{r convergence}
plot(resultsPumpAdam$param, type="l",
     main="pump", xlab=expression(alpha), ylab=expression(beta))
```

The default option is Adam. One can change the method using the argument `method=...`.
For example, suppose we want to apply the 1D sampling method. We set `method="ga1D`.

```{r algorithm, results="hide"}
resultsPump1Dsampling <- computeMLE(model=pump, 
                                    paramNodes = paramNodesPump, 
                                    paramInit=c(10, 10),
                                    compiledFuns=compiledFunsPump,
                                    method="ga1D")
```

The result agrees with what we get from Adam:
```{r 1DResult}
resultsPump1Dsampling$MLE
```
