---
title: "glmmExample"
author: "Sara Stoudt"
date: "April 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generalized Linear Mixed Model (GLMM) example

Now we will explore a real-world example and show how to change some of the more nitty-gritty details. These are  

```{r}
library("MCMCmaxlik")
```

Here we set up the salamander data.

```{r}
require(glmm)
data(salamander)
names(salamander)
nrow(salamander) 

isRR=ifelse(salamander$Cross=="R/R",1,0)
#head(isRR)

isRW=ifelse(salamander$Cross=="R/W",1,0)
#head(isRW)

isWR=ifelse(salamander$Cross=="W/R",1,0)
#head(isWR,20)

isWW=ifelse(salamander$Cross=="W/W",1,0)
#head(isWW,20)


```

Here we set up the model.

```{r model}
glmmCode <- nimbleCode({
  for(i in 1:N){
  logit(theta[i])<-beta1*isRR[i]+beta2*isRW[i]+beta3*isWR[i]+beta4*isWW[i]+REF[i]+REM[i]
  y[i]~dbin(theta[i],1)
  REF[i]~dnorm(0,varF)
  REM[i]~dnorm(0,varM)
  }
  
  
  beta1 ~ dnorm(0, sd = 10000)
  beta2 ~ dnorm(0, sd = 10000)
  beta3 ~ dnorm(0, sd = 10000)
  beta4 ~ dnorm(0, sd = 10000)
  varF ~ dunif(0, 1000)
  varM ~ dunif(0, 1000)
  
})


glmmConstants <- list(N = nrow(salamander))

glmmData <- list(
  isRR=isRR,isWR=isWR,isRW=isRW,isWW=isWW,y=salamander$Mate
)

glmmInits <- list(beta1 = 1, beta2 = 1, beta3=1,beta4=1,varF=5,varM=5)

glmmMod <- nimbleModel(code=glmmCode, constants=glmmConstants, data=glmmData, inits=glmmInits, check = FALSE)

glmmParamNodes <- glmmMod$getNodeNames(topOnly = T,stochOnly=T)

Cglmm <- compileNimble(glmmMod)

```


## Running the algorithms

First, we compile the algorithms in NIMBLE using `buildMCMCmaxlik()`.

```{r compile}
compiledFunsGlmm <- buildMCMCmaxlik(glmmMod, glmmParamNodes)
```

Then we can run the algorithms using `computeMLE()`. For Adadelta and Adam we use the default settings used by the creators of these step size schedules, but we do allow the user to change these as desired. We want to give flexibility for those who have experience with these algorithms in other contexts and may want to customize our implementations to fit their needs. 

Default Parameters for Adadelta: delta = 1e-04, eps = 1e-2

Default Parameters for Adam: beta1=0.9, beta2=0.999

```{r algorithm1D, results="hide"}
resultsGLMMadadeltaDefault <- computeMLE(model=glmmMod, 
                                    paramNodes = glmmParamNodes, 
                                    paramInit=rep(2, 6),
                                    compiledFuns=compiledFunsGlmm,
                                    method="adadelta")

```


```{r algorithm2D, results="hide"}
resultsGLMMadamDefault<- computeMLE(model=glmmMod, 
                                    paramNodes = glmmParamNodes, 
                                    paramInit=rep(2, 6),
                                    compiledFuns=compiledFunsGlmm,
                                    method="adam")


```

Now let's change some parameters.

```{r algorithm1, results="hide"}
resultsGLMMadadelta <- computeMLE(model=glmmMod, 
                                    paramNodes = glmmParamNodes, 
                                    paramInit=rep(2, 6),
                                    compiledFuns=compiledFunsGlmm,
                                    method="adadelta",delta=5e-04,eps=.1)


```


```{r algorithm2, results="hide"}
resultsGLMMadam<- computeMLE(model=glmmMod, 
                                    paramNodes = glmmParamNodes, 
                                    paramInit=rep(2, 6),
                                    compiledFuns=compiledFunsGlmm,
                                    method="adam",beta1=.8,beta2=0.99)


```

How do we do? We can see that the default parameters yield more consistent results between AdaDelta and Adam, and that the results for this example are sensitive to the tuning parameters of these methods. The default values we implement are those used by the creators of AdaDelta and Adam, and we recommend sticking with these.

```{r 1DResult}
resultsGLMMadadeltaDefault$MLE
resultsGLMMadamDefault$MLE
resultsGLMMadadelta$MLE
resultsGLMMadam$MLE
```

